Future cosmic microwave background delensing with galaxy surveys

### 2018-01-22 Notes/referee_response.tex

\documentclass[12pt]{amsart}
\usepackage[top=2. cm, bottom=2. cm, left=1.0in, right=1.0in]{geometry}
%\geometry{a4paper} % or letter or a5paper or ... etc
% \geometry{landscape} % rotated page geometry
\usepackage[applemac]{inputenc}
\usepackage{amsmath,amssymb} 
% See the ``Article customise'' template for come common customisations
\usepackage{bm}
\usepackage{epsfig}
\usepackage{graphicx,epsfig}
\usepackage{hyperref}
\usepackage{ifthen}
\usepackage [english]{babel}
\usepackage{xstring}
\usepackage[applemac]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{color}
\usepackage{epstopdf}
\usepackage{caption}
\captionsetup{justification=centerlast,singlelinecheck=false,font=small,skip=2pt}
% notes to make comment is non-zero if
\newcommand{\al}[1]{\begin{align} #1 \end{align}}

@@

\title{  \scriptsize{Response to Referee for: "Future cosmic microwave background delensing with galaxy surveys"}}
\author{A. Manzotti}
%\date{} % delete this line to display the current date

%%% BEGIN DOCUMENT
\begin{document}
\pagestyle{empty}
\date{\today}
\maketitle
%\date{\today}
I want to thank the referee for the careful review of the paper, and for the comments, corrections, and suggestions that ensued.

\vspace{0.1cm}

Below, we sequentially address all of the points raised.

\vspace{0.1cm}

While working on this answer, I found a minor bug in the part of the code that plots the CIB kernel in Fig 1 (the CIB kernel peak is then slightly shifted). This is fixed now, and it did not affect the results. Also after working on point 5 below, I switched the SKA kernel in Fig. 1 from the optimistic 0.1 $\mu$Jy to the more realistic 10 $\mu$Jy casee. Furthermore, I explicitly refer to that in the label. The 0.1 $\mu$Jy threshold case was discussed in an earlier version of this paper and it was mistakenly left in the Figure. I am sorry for the confusion.

\vspace{1.cm}

@@

\begin{enumerate}
\item The abstract is unnecessarily very long and could be considerably
shortened.

\vspace{0.4cm}
I agree.
The abstract length is now reduced to $\sim 80\%$ of its previous length, mainly removing redundancy and quantitative results that can be found easily in the tables. I believe it is still a little bit longer than the average but reasonably so.
\vspace{1cm}

@@

\item Putting the lmin-lmax ranges in the tables as well as having the
dashed grey line of Fig. 5 onto Fig 6-7-8 would help the readability
of the paper. 
\vspace{0.4cm}

I agree with the referee on this point. I first thought this could cause a bit of clatter, but the gain in readability is worth the risk.
I added the dashed line to all the figures. The $\ell$ range used for the inflationary constraints is also at the end of each table both for readability and reproducibility. The choice was quite arbitrary tough. We used all the scales available to the instrument. The range where the bulk of the signal to noise is, has already been explored in some of the referenced work (\cite{dodelson:2014,abazajian:2016} etc.).

@@

We agree that the phrasing of the sentences where secondary B-modes sources are introduced can be confusing. While it is well known in the field, it is important to be clear to a broader audience.
For this reason, we add this sentence in the introduction:

\vspace{1cm}
\textit{Two main effects produce non-primordial B-modes: the polarized foregrounds from the Galaxy  and the effect of gravitational interactions with large scale structures (LSS). 
These two components need to be treated with very different techniques and in this work, we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).}

\vspace{0.2 cm}

Note that if the referee meant that the galactic B-modes component is the only relevant secondary component compared to lensing B-modes, I do not think that is necessarily true. 
While on most of the sky the galactic component is the higher contribution in power, that is not true for an opportunely chose small (say 1\%) patch. 
This is the size we plan for future experiments like CMB-S4.
Furthermore, that is probably not the best metric.
As shown in the CMB-S4 science book \cite{abazajian:2016} if one wants to maximize the $r$ constraints, lensing and foregrounds are equally important. 
The importance there is measured as the number of detectors used to either improve the lensing reconstruction or to observe the sky at different frequencies to clean the foregrounds component.
This depends slightly on the fraction of the sky used, but it is true for a wide range that contains all possible future ground experiments.

@@

\vspace{1cm}
\item The author introduces sharp cuts to CIB and WISE at $\ell < 100$. Is the
alternative approach of incorporating the foreground content in the
auto spectra of the tracer (E.g. Eq. 17) combination possible ? That
would retain some more of the signal presumably. 
\vspace{0.4cm}

This is indeed a good idea. The real obstacle here is our current knowledge of foreground spectra. There is not an accurate model for that (at higher scales we use a simple power law). However one might measure that empirically, given a dataset with enough signal to noise on those large scales, by measuring $C^{II}$ and $C^{I\phi}$ directly from the data. Doing that will call everything in the tracer map that does not correlate with $\phi$, "noise". That noise will include foregrounds too. However, it is unclear if eventual dust residual can cause biases in the delensed B-modes (if they correlate with dust contamination in CMB polarizations). This is certainly small, but maybe at the level of very futuristic CMB inflationary constraints.
Given the small amount of delensing power for $\ell<100$ the field decided, so far, to cut those in forecasts. Foregrounds are quite complicated and therefore I believed only data would answer this. An exploration of these on simulations is probably beyond the scope of this paper, but very soon future delensing work will test this on real data.

\vspace{0.3cm}

I added added a sentence to the paper regarding this possible way forward.

@@

\vspace{1cm}
\item Fig 1 might look confusing at first. The SKA kernel there shows
brilliant overlap with the CMB lensing kernel, but is not the most
powerful tracer. On the other hand, DES for instance has very moderate
overlap but can still achieve very high performance on large scales.
Can the author comment and help the reader make sense of that figure ?

As pointed out at the very beginning the confusion was probably caused by a mistake: I was plotting an optimistic SKA case (with a threshold at $0.1\mu$Jy) instead of the
$10\mu$Jy case discussed in the paper. Again, thanks for catching this.

\vspace{0.2cm}

However I agree that Fig 1 may be a little confusing because it does not tell the entire story. 
The redshift overlap is not the only ingredient for the delensing efficiency. For example, the noise is relevant as well, and this is why for example for S2 experiments CIB can be better than the CMB internal reconstruction even if the latter has a perfect overlap. 
Additionally, not all the redshifts matter at the same level since they correspond to different angular scales (depending on the distance of the tracer) which, as shown by the dashed curves in the figures 5-8, contributes differently to the generation of CMB B-modes.
Showing all of this in one plot is hard.

However, I think Fig 1 tells most of the story. For example, DES is quite less efficient than SKA in delensing efficiency (16\%) vs (52\%) and SKA is one of the most efficient LSS survey.
Furthermore, I think this figure highlights the central part (the "kernel") that characterize the calculation of different tracers spectra (using Eq. 2) for people not familiar with it.

@@

\item Is it clear, as the author claims in IV A with no supporting
evidence, that fixing the cosmology has no impact on the {r,nt}
constraints ? The cosmology fixes the residual B-mode amplitude, which
is unequally important in the delensed vs nominal cases. An slightly
improved discussion would be welcome here.

\vspace{0.4cm}
Note that indeed, as specified in the text, the chosen cosmology \textit{has} an impact on the absolute value of the ($r$,$n_{t}$) constraints.
What I claim cosmology has almost no impact on, in particular if compared to other uncertainties, is the delensing efficiency (how much these constraints improve), i.e. the ratio $C^{BB}_{\ell}$ nominal/delensed. What matters for the efficiency is the \textit{fraction} of removed power. That depends only on the factor  $1-\rho^2$ where $\rho$ is the correlation factor. 
The correlation factor has a very little dependence on cosmology in particular given the tight constraints from Planck on $\Omega_{m}h^2$ and $\sigma_{8}$, and it is expected to be a secondary uncertainty here compared to, for example, the redshift distribution of tracers. Indeed $\rho$ contains $C^{II}$ $C^{I\phi}$, $C^{\phi\phi}$ that are integrals, with slightly different weights in redshift, of the cosmology dependent power spectrum $P(k)$. For this reason, they should scale similarly with cosmology leading to a substantial cancellation in the cosmological dependence of $\rho$.

Different cosmologies can also shifts the scales where the information on ($r$, $n_{t}$) come from by changing the shape of the $C_{\ell}^{BB}$ spectra.

I agree that testing this intuition is important. For this reason, I run the same pipeline with $10\%$ higher $\Omega h^{2}$, which is almost 40 times more than what Planck constraints \citep{planck-collaboration:2016b}. The efficiency values changes, but by at most $8-9\%$. Given that the Planck constraints are way tighter I consider this, as expected, a present but subdominant effect. 
For example for $1\%$ higher $\Omega h^{2}$ the efficiency values changes by at most $1.5\%$.

I added few sentences to the paper to report this test.

@@

\item The author makes a strong point that survey data should be split
into bins, as this makes the correlation increases by 30\%. I am not
entirely convinced that this is as relevant as claimed. From Fig 4, it
seems that the improvement is fairly limited on the scales that are
relevant for B-mode delensing. Presumably, that is more important for
T or E delensing. The author might want to comment or adapt.

\vspace{0.4cm}
In general, I agree with the point that quoting the improvement in the correlation is not entirely fair because the scales at which that improvement happens is also necessary to understand how much power will be removed.
For this reason, I added in the abstract this quantity too.

Furthermore, while I agree that the improvement would be even more significant for T-E delensing, it is nonetheless also quite important for BB too.
As can be seen in Figure 4 in the paper LSST shows a significant improvement in the correlation with the lensing potential all the way to $\ell\sim 800$. Even for other probes, in general, less powerful, there is still an improvement for example at \textit{all} the scales where the contribution to BB (the dashed line) is relevant $50<\ell<400$.
Below I am attaching a table that shows the improvement in the BB power removed and constraints on $\sigma(r)$.
Note that the 30\% level quoted in the abstract corresponds to the most futuristic tracer, i.e., LSST. That should be explicitly stated, so I added this to the abstract.

\begin{table}[htp]
\caption{Improvement on $\sigma(r)$ constraints together with the BB power removed (in parenthesis) with and without binning. Note that the improvement is noise-dependent but the removed power is not.}
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
Survey & No Binning & Binning\\
\hline
DES & 1.17 (14\%)& 1.20 (16\%) \\
\hline
LSST & 1.67 (40\%) & 2.10 (52\%)\\
\hline
DESI & 1.10 (9\%) & 1.13 (12\%)  \\
\hline
\end{tabular}
\end{center}
\label{default}
\end{table}%

### 2018-01-22 Notes/galaxy_delensing.tex
journal version

The primordial B-modes component of the cosmic microwave background (CMB) polarization is a promising experimental dataset to probe the inflationary paradigm. B-modes are indeed a direct consequence of the presence of gravitational waves in the early universe. However, several secondary effects in the low redshift universe will produce \textit{non-primordial} B-modes. In particular, the gravitational interactions of CMB photons with large-scale structures will distort the primordial E-modes, adding a lensing B-mode component to the primordial signal. Removing the lensing component ("delensing") will then be necessary to constrain the amplitude of the primordial gravitational waves. Here we discuss the role of current and future large-scale structure surveys in a multi-tracers approach to CMB delensing. We find that, in general, galaxy surveys should be split into tomographic bins as this can increase the reduction of lensing B-modes by $\sim 25\%$ in power in the most futuristic case. 
Ongoing or recently completed CMB experiments (CMB-S2) will particularly benefit from large-scale structure tracers that, once properly combined, will have a better performance than a CMB internal reconstruction. With the decrease of instrumental noise, the amount of lensing B-modes power removed using CMB internal reconstruction alone will rapidly increase. Nevertheless, optical galaxy surveys will still play an important role even for CMB S4. In particular, an LSST-like survey can a achieve a delensing performance comparable to a 3G CMB experiment but with entirely different systematics. This redundancy will be essential to prove the robustness against systematics of an eventual detection of primordial B-modes.

@@

Because the CMB B-modes provide the cleanest known observational window into the primordial gravitational wave background, improving their measurement is a major objective of current and future CMB experiments.
Even if inflationary B-modes have not been detected yet, the target value $r \gtrsim 10^{-3}$, should be reachable shortly given the level of noise expected in future CMB experiments \citep{kamionkowski:2016,abazajian:2016}.
However, just reducing the level of noise will not be enough to reach this target.
Indeed, the observed B modes are not solely sourced by early universe physics; they are also produced by secondary effects taking place in the late-time low-redshift universe.
\textcolor{blue}{Two main effects produce non-primordial B-modes: the polarized foregrounds from the Galaxy  and the effect of gravitational interactions with large scale structures (LSS). 
These two components need to be treated with very different techniques, and in this work, we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).}
Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
This expected component has now been measured both in cross-correlation with LSS \citep{hanson13,bicep2-collaboration:2016,polarbear2014c,van-engelen:2015,planck2015XV} and from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.

@@

The CMB lensing potential is \textit{the} field that we need in order to reverse the effect of large scale structure and delens the CMB.
However, the lensing potential can also be reconstructed using the CMB itself. In that case, we can treat it as a noisy tracer of the true field. Both the CMB lensing field and its noisy reconstructed counterpart have the same kernel $W^{\kappa}(z)$.
However, when computing the power spectrum of the latter, we need to add a noise component. Given the instrumental noise level and the beam, we can calculate the reconstruction noise $N_{\ell}^{\kappa \kappa}$, and so
\beq\label{eqn:noisekappa}
\cl^{\kappa_\mathrm{rec} \kappa_\mathrm{rec}} = \cl^{\kappa \kappa } +N^{\kappa \kappa}_{\ell}
\eeq
In this work, the level of noise is computed assuming an iterative approach to the CMB lensing reconstruction as described in \cite{smith:2012,hirata:2003}. \textcolor{blue}{The iterative approach has not been sufficiently demonstrated on data yet, but it is among the future goals of the CMB community \cite{abazajian:2016}.
Note however that, as pointed out in \cite{smith:2012,hirata:2003}, this improves the CMB lensing reconstruction compared to a quadratic estimator approach mainly for S3 and S4 noise levels. For these two cases, we also show the delensing efficiency obtained when the CMB lensing is reconstructed with the standard quadratic estimator (QE) approach \cite{okamoto:2003,hu:2002}.
Furthermore, we do not deal here with the presence of an internal delensing bias. Several promising approaches have been proposed \cite{sehgal:2017,namikawa:2017} even if they have not been applied to low noise data yet \cite{carron:2017}.}

@@

% ===================
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./images/compare_kernel.pdf}
\caption{Kernel Comparison: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis. The larger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential is thus leading to a higher delensing efficiency. 
\textcolor{blue}{However, the efficiency is not exactly proportional to the overlapping area. 
Firstly structure at different redshifts contribute differently to the generation of large-scale B-modes because of the geometric properties of the lensing kernel, and secondly, we are not taking the tracers noise into account here.
Note that for galaxy surveys the kernel corresponds to $b(z)\frac{dN}{dz}$ and not just the redshift distribution of galaxies.}}
\label{fig:kernel}
\end{center}
\end{figure}
% ===================

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./images/BB_res_ell2.pdf}
\caption{Here we illustrate the effect of delensing on the B-mode power spectrum. The orange solid line corresponds to the fiducial lensing B-mode component of the signal while the dashed blue line corresponds to the inflationary one for the higher amplitude allowed by current experiments.
These are compared with the residual power left after delensing using some of the LSS tracers described in \refsec{model}. 
The rapid improvement in the level of instrumental noise (dashed curves for SPTPol and CMB S4) will require a high delensing efficiency to exploit these experiments fully.}
\label{fig:bbplot}
\end{center}
\end{figure}
% ===================

@@

% ===================
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./images/compare_kernel_tomo.pdf}
\caption{Comparison of kernels of the 10 LSST tomographic bins together with the full LSST survey and the CMB lensing kernel. 
Compared to a full survey approach, tomographic binning allows to optimally weight different bins according to their cross-correlation with the CMB lensing. This leads to a better delensing efficiency. }
\label{fig:kerneltomo}
\end{center}
\end{figure}
% ===================

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./images/B_res_bin.pdf}
\caption{Tomographic bins improve the cross-correlation of galaxy surveys with CMB lensing. Here we show the cross-correlation coefficient (\refeq{rho1}) as a function of angular scale for full surveys (solid lines) and tomographically binned surveys (dot-dashed line). For details about the binning, see \refsec{tomo-bin} and in particular \refeq{zbin}. }
\label{fig:rho-tomo}
\end{center}
\end{figure}

@@

In \refeq{fisher} we are making a few important assumptions. 
First, we are fixing all the cosmological parameters except $\{r,n_{t}\}$. Uncertainties in those will propagate to larger uncertainties in $\{r,n_{t}\}$. While neglecting this will lead to slightly optimistic absolute constraints, it has no significant impact on the estimate of the improvement due to delensing. \textcolor{blue}{For example, running the same pipeline with $10\%$ higher $\Omega h^{2}$, which is almost 40 times more than what Planck constraints \cite{planck-collaboration:2016b}, change the delensing improvement factors by at most $7-8\%$. A more reasonable 1\% higher $\Omega h^{2}$ corresponds to variations in the improvement factors of at most 1.5\%.}

@@

On the CMB side, we will combine a BICEP-Keck like \textit{deep} CMB experiment with an overlapping \textit{higher resolution} experiment. For the deep experiment we assume an instrumental noise equal to $3\mu$K-arcmin in polarization ($\sqrt{2}$ lower in temperature), a beam of 30 arcmin and an angular scale range of $50<\ell<500$.
\textcolor{blue}{Here and in the following sections, we use all scales available to the instrument, even if the bulk of the signal to noise is at large scales $\ell<500$ \cite{dodelson:2014}.}
We assume that the CMB lensing reconstruction is performed by the higher resolution CMB experiment.
We explore two possibilities.
First, we test the delensing efficiency with an internal reconstruction of CMB lensing performed by the Planck satellite \cite{planck-collaboration:2016a}. In this case, for the noise in the CMB lensing map in \refeq{noisekappa} we use the actual noise curves publicly available
\footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}. For consistency, we also combine a Planck-like E-mode map of $60\mu$K-arcmin in polarization with a beam equal to 7 arcmin that however adds almost nothing to the \textit{deep} CMB experiment.
Then, we test a second scenario with a ground-based experiment characterized by noise levels in polarization consistent with SPT-Pol \cite{henning:2017}: $9.4\mu$K-arcmin in polarization with a beam equal to 1.2 arcmin. In this case we set the angular scale range at $50<\ell<3000$. However, the results are quite robust against this choice.

We combine these CMB experiments with the CIB and current low redshift galaxy surveys like DES and WISE and we compute the improvement in the delensing efficiency.
Following \cite{yu:2017} we cut both the CIB and WISE at $\ell<100$ where they are contaminated by large Galactic dust residuals. 
\textcolor{blue}{In the future, with better theoretical foreground models or high signal to noise measurements, we might be able to characterize these fields (and their spectra $C_{\ell}^{\kappa I}$ and $C_{\ell}^{II}$) even at larger scales. Potential biases effect may be present like, for example, the correlation between the foreground contamination in galaxies and in the CMB polarization itself. 
If these are under control, we might then be able to also use the $\ell<100$ component of these tracers to delens. However since the lensing contribution to B-modes from LSS modes at these scales is quite small, there is not too much to gain in term of delensing efficiency.}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./images/actual_scenario_planck.pdf}
\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$ for completed and ongoing Stage-2 tracers.
Here we show both current galaxy survey and internally reconstructed CMB lensing potential. 
The dashed curve highlights which scales contribute to the $\ell<100$ B-mode power and are thus most useful to delens. It corresponds to (with arbitrary scale) $C^{\kappa \kappa} \times \frac{\partial C^{BB}}{\partial{C^{\kappa \kappa}}}$}
\label{fig:corrnow-planck}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./images/actual_scenario.pdf}
\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$ for completed and ongoing Stage-2 tracers. 
This figure is the same as \reffig{corrnow-planck} but using SPTPol instead of Planck as the high-res CMB experiment performing the internal reconstruction.
}
\label{fig:corrnow}
\end{center}
\end{figure}

@@

The correlation factor attainable using generation 3 experiments is shown in \reffig{corrS3}.
An interesting finding is that DESI will be less efficient (removing $11\%$ of the power) than a DES-like survey ($17\%$) despite the fact that it can probe slightly higher redshift. The reason is that, because of the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES increases the delensing efficiency. Unfortunately, adding DESI will only bring the power removed using LSS from 41 to 45\%.
Contrary to current S2 experiments, the CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 600$.
\textcolor{blue}{For CMB S3, the improvement due to an iterative approach compared to the quadratic estimator is present but not significant, so that a more traditional technique can be used. For example, in this case, the removed power goes from $56\%$ to $51\%$ (quoted in \reftab{S3}) when using a quadratic estimator (QE) instead of the iterative one.}

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./images/S3_scenario.pdf}
\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$. This figure is the same as \reffig{corrnow} but with a stage-3 CMB experiment performing the internal reconstruction and with DESI added to the stage-2 galaxy surveys (LSS-S2). }
\label{fig:corrS3}
\end{center}
\end{figure}

@@

The correlation factor with CMB lensing for Stage-4 experiments is shown in \reffig{corrs4_10}.
%=======
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./images/S4_scenario_ska10.pdf}
\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$. This figure is the same as \reffig{corrnow} but with a stage-4 CMB experiment performing the internal reconstruction and with LSST and SKA added to the stage-2 and stage-3 galaxy surveys (LSS-S3). }
\label{fig:corrs4_10}
\end{center}
\end{figure}

@@

% ===================
\begin{figure*}[ht]
\includegraphics[scale=0.8]{./images/errors_summary2.pdf}
\caption[]{A summary of the amount of lensing B-modes power removed by different tracers for different generations of experiments.

>>>>

The primordial B-modes component of the cosmic microwave background (CMB) polarization is a promising experimental dataset to probe the inflationary paradigm. B-modes are indeed a direct consequence of the presence of gravitational waves in the early universe. However, several secondary effects in the low redshift universe will produce \textit{non-primordial} B-modes. In particular, the gravitational interactions of CMB photons with large-scale structures will distort the primordial E-modes, adding a lensing B-mode component to the primordial signal. Removing the lensing component ("delensing") will then be necessary to constrain the amplitude of the primordial gravitational waves. Here we examine the role of current and future large-scale structure surveys in a multi-tracers approach to CMB delensing. We find that, in general, galaxy surveys should be split into tomographic bins as this can increase the reduction of lensing B-modes by $\sim 25\%$ in power in the most futuristic case. 
Ongoing or recently completed CMB experiments (CMB-S2) will particularly benefit from large-scale structure tracers that, once properly combined, will have a better performance than a CMB internal reconstruction. With the decrease of instrumental noise, the lensing B-modes power removed using CMB internal reconstruction alone will rapidly increase. Nevertheless, optical galaxy surveys will still play an important role even for CMB S4. In particular, an LSST-like survey can a achieve a delensing performance comparable to a 3G CMB experiment but with entirely different systematics. This redundancy will be essential to demonstrate the robustness against systematics of an eventual detection of primordial B-modes.

@@

Because the CMB B-modes provide the cleanest known observational window into the primordial gravitational wave background, improving their measurement is a major objective of current and future CMB experiments.
Even if inflationary B-modes have not been detected yet, the target value $r \gtrsim 10^{-3}$, should be reachable shortly given the level of noise expected in future CMB experiments \citep{kamionkowski:2016,abazajian:2016}.
However, just reducing the level of noise will not be enough to reach this target.
Indeed, the observed B modes are not solely sourced by early universe physics; they are also produced by secondary effects taking place in the late-time low-redshift universe.
Two main effects produce non-primordial B-modes: the polarized foregrounds emission from the Galaxy and the gravitational interactions of the primordial CMB with large scale structures (LSS). 
These two components need to be treated with different techniques, and in this work, we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).
Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
This expected component has now been measured both in cross-correlation with LSS \citep{hanson13,bicep2-collaboration:2016,polarbear2014c,van-engelen:2015,planck2015XV} and from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.

@@

The CMB lensing potential is \textit{the} field that we need in order to reverse the effect of large scale structure and delens the CMB.
However, the lensing potential can also be reconstructed using the CMB itself. In that case, we can treat it as a noisy tracer of the true field. Both the CMB lensing field and its noisy reconstructed counterpart have the same kernel $W^{\kappa}(z)$.
However, when computing the power spectrum of the latter, we need to add a noise component. Given the instrumental noise level and the beam, we can calculate the reconstruction noise $N_{\ell}^{\kappa \kappa}$, and so
\beq\label{eqn:noisekappa}
\cl^{\kappa_\mathrm{rec} \kappa_\mathrm{rec}} = \cl^{\kappa \kappa } +N^{\kappa \kappa}_{\ell}
\eeq
In this work, the level of noise is computed assuming an iterative approach to the CMB lensing reconstruction as described in \cite{smith:2012,hirata:2003}. 
Note that, as pointed out for example in \cite{smith:2012,hirata:2003}, this iterative approach improves the CMB lensing reconstruction if compared to a quadratic estimator (QE) approach \cite{okamoto:2003,hu:2002} for S3 and S4 noise levels. For these two cases, we also compute the delensing efficiency obtained when the CMB lensing is reconstructed with the standard quadratic estimator (QE).
Furthermore, we do not deal here with the presence of an internal delensing bias. Indeed we assume that we can use all the CMB scales for lensing reconstruction even if these scales are also used to constraint inflationary parameters. These is supported by several promising approaches \cite{sehgal:2017,namikawa:2017} although they have not been applied to low noise data yet (see \cite{carron:2017} as an example where the internal delensing bias was removed).

@@

% ===================
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./compare_kernel.pdf}
\caption{Kernel Comparison: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis. The larger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential is thus leading to a higher delensing efficiency. However, the efficiency is not exactly proportional to the overlapping area. 
Firstly, structures at different redshifts contribute differently to the generation of large-scale B-modes because of the geometric properties of the lensing kernel, and secondly, we are not taking the tracers noise into account here.
Note that for galaxy surveys the kernel corresponds to $b(z)\frac{dN}{dz}$ and not just the redshift distribution of galaxies.}
\label{fig:kernel}
\end{center}
\end{figure}
% ===================

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./BB_res_ell2.pdf}
\caption{Here we illustrate the effect of delensing on the B-mode power spectrum. The orange solid line corresponds to the fiducial lensing B-mode component of the signal while the dashed blue line corresponds to the inflationary one for the higher amplitude allowed by current experiments.
These are compared with the residual power left after delensing using some of the LSS tracers described in \refsec{model}. 
The rapid improvement in the level of instrumental noise (dashed curves for SPTPol and CMB S4) will require a high delensing efficiency to exploit these experiments fully.}
\label{fig:bbplot}
\end{center}
\end{figure}
% ===================

@@

% ===================
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./compare_kernel_tomo.pdf}
\caption{Comparison of kernels of the 10 LSST tomographic bins together with the full LSST survey and the CMB lensing kernel. 
Compared to a full survey approach, tomographic binning allows to optimally weight different bins according to their cross-correlation with the CMB lensing. This leads to a better delensing efficiency. }
\label{fig:kerneltomo}
\end{center}
\end{figure}
% ===================

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./B_res_bin.pdf}
\caption{Tomographic bins improve the cross-correlation of galaxy surveys with CMB lensing. Here we show the cross-correlation coefficient (\refeq{rho1}) as a function of angular scale for full surveys (solid lines) and tomographically binned surveys (dot-dashed line). For details about the binning, see \refsec{tomo-bin} and in particular \refeq{zbin}. }
\label{fig:rho-tomo}
\end{center}
\end{figure}

@@

In \refeq{fisher} we are making a few important assumptions. 
First, we are fixing all the cosmological parameters except $\{r,n_{t}\}$. Uncertainties in those will propagate to larger uncertainties in $\{r,n_{t}\}$. While neglecting this will lead to slightly optimistic absolute constraints, it has no significant impact on the estimate of the improvement due to delensing. For example, running the same pipeline with $10\%$ higher $\Omega h^{2}$, which is almost 40 times more than what Planck constraints allow \cite{planck-collaboration:2016b}, change the delensing improvement factors by at most $7-8\%$. A more reasonable 1\% higher $\Omega h^{2}$ corresponds to variations in the improvement factors of at most 1.5\%.

@@

On the CMB side, we will combine a BICEP-Keck like \textit{deep} CMB experiment with an overlapping \textit{higher resolution} experiment. For the deep experiment we assume an instrumental noise equal to $3\mu$K-arcmin in polarization ($\sqrt{2}$ lower in temperature), a beam of 30 arcmin and an angular scale range of $50<\ell<500$.
Here and in the following sections, we use all the scales measured by the instrument, even if the bulk of the signal to noise is at large scales $\ell<500$ \cite{dodelson:2014}.
We assume that the CMB lensing reconstruction is performed by the higher resolution CMB experiment.
We explore two possibilities.
First, we test the delensing efficiency with an internal reconstruction of CMB lensing performed by the Planck satellite \cite{planck-collaboration:2016a}. In this case, for the noise in the CMB lensing map in \refeq{noisekappa} we use the actual noise curves publicly available
\footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}. For consistency, we also combine a Planck-like E-mode map of $60\mu$K-arcmin in polarization with a beam equal to 7 arcmin that however adds almost nothing to the \textit{deep} CMB experiment.
Then, we test a second scenario with a ground-based experiment characterized by noise levels in polarization consistent with SPT-Pol \cite{henning:2017}: $9.4\mu$K-arcmin in polarization with a beam equal to 1.2 arcmin. In this case we set the angular scale range at $50<\ell<3000$. However, the results are quite robust against this choice.

We combine these CMB experiments with the CIB and current low redshift galaxy surveys like DES and WISE and we compute the improvement in the delensing efficiency.
Following \cite{yu:2017} we cut both the CIB and WISE at $\ell<100$ where they are contaminated by large Galactic dust residuals. 
In the future, with better theoretical foreground models or high signal to noise measurements, we might be able to characterize these fields (and their spectra $C_{\ell}^{\kappa I}$ and $C_{\ell}^{II}$) even at larger scales. Potential biases effect may be present like, for example, the correlation between the foreground contamination in galaxies and in the CMB polarization itself. 
However, we expect these biases to be small, and we might then be able to also use the $\ell<100$ component of these tracers to delens. However since the lensing contribution to B-modes from LSS modes at these scales is quite small, there is not too much to gain in term of delensing efficiency.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./actual_scenario_planck.pdf}
\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$ for completed and ongoing Stage-2 tracers.
Here we show both current galaxy survey and internally reconstructed CMB lensing potential. 
The dashed curve highlights which scales contribute to the $\ell<100$ B-mode power and are thus most useful to delens. It corresponds to (with arbitrary scale) $C^{\kappa \kappa} \times \frac{\partial C^{BB}}{\partial{C^{\kappa \kappa}}}$}
\label{fig:corrnow-planck}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./actual_scenario.pdf}
\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$ for completed and ongoing Stage-2 tracers. 
This figure is the same as \reffig{corrnow-planck} but using SPTPol instead of Planck as the high-res CMB experiment performing the internal reconstruction.
}
\label{fig:corrnow}
\end{center}
\end{figure}

@@

The correlation factor attainable using generation 3 experiments is shown in \reffig{corrS3}.
An interesting finding is that DESI will be less efficient (removing $11\%$ of the power) than a DES-like survey ($17\%$) despite the fact that it can probe slightly higher redshift. The reason is that, because of the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES increases the delensing efficiency. Unfortunately, adding DESI will only bring the power removed using LSS from 41 to 45\%.
Contrary to current S2 experiments, the CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 600$.
For CMB S3, an iterative approach will improve the CMB lensing reconstruction compared to the quadratic estimator, even if not very significantly. For example, in this case, the removed power goes from $56\%$ to $51\%$ (quoted in \reftab{S3}) when using a quadratic estimator (QE) instead of the iterative one.

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./S3_scenario.pdf}
\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$. This figure is the same as \reffig{corrnow} but with a stage-3 CMB experiment performing the internal reconstruction and with DESI added to the stage-2 galaxy surveys (LSS-S2). }
\label{fig:corrS3}
\end{center}
\end{figure}

@@

The correlation factor with CMB lensing for Stage-4 experiments is shown in \reffig{corrs4_10}.
%=======
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./S4_scenario_ska10.pdf}
\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$. This figure is the same as \reffig{corrnow} but with a stage-4 CMB experiment performing the internal reconstruction and with LSST and SKA added to the stage-2 and stage-3 galaxy surveys (LSS-S3). }
\label{fig:corrs4_10}
\end{center}
\end{figure}

@@

% ===================
\begin{figure*}[ht]
\includegraphics[scale=0.8]{./errors_summary2.pdf}
\caption[]{A summary of the amount of lensing B-modes power removed by different tracers for different generations of experiments.
